Deciding Bitvector Termination with Program Synthesis

### 2014-07-04 papers/termination/paper.tex
Contribution
Move related work to end
Added theorems

@@

\section{Related Work}
%\subsection{Termination analysis}
Automated program termination is a research topic that has received a fair amount of attention from the software verification community.
In order to compare our technique to the rest of the area, 
Figure~\ref{fig:handletable} summarises the related works with respect to the restrictions they impose on the transition relations as well as the form of the computed ranking functions.

@@

\noindent {\bf Unconditional termination.}
To show that $L$ is unconditionally terminating (i.e. it eventually terminates regardless of the
state it starts in), we can find a ranking function for $L$.
The existence of a ranking function is equivalent to the satisfiability
of the formula from Definition~\ref{def:UT}, so a witness to the satisfiability
of formula {\bf [UT]} is also a ranking function and thus a proof of $L$'s unconditional termination.

@@

@@

@@

Conditional termination is then equivalent to the
formula {\bf [CT]} of Defintion~\ref{def:ct}. If this formula is satisfiable,
two witnesses are returned:
\begin{itemize}
\item $W$ is an inductive invariant of $L$ that is established by the initial states $I$ if the loop
guard $G$ is met.
\item $R$ is a ranking function for $L$ as restricted by $W$ -- that is to say, $R$ need only
be well founded on those states satisfying $W \wedge G$.  Since $W$ is an inductive invariant of $L$,
$R$ is strong enough to show that $L$ terminates from any of its initial states.
\end{itemize}

@@

\begin{proof}
 {\sc Headshot}'s output constitutes a constructive, machine checked proof of (non-)termination.
\end{proof}

@@

>>>>

\item Our procedure is based on a series of source-to-source transformations.  As well as leading to a concise implementation,
this has the side effect of ensuring that our proof is correct for the exact binary produced by a particular compiler,
rather than for an abstraction of the program.

@@

@@

\noindent {\bf Unconditional termination.}
To show that $L$ is unconditionally terminating (i.e. it eventually terminates regardless of the
state it starts in), we can find a ranking function for $L$.

\begin{theorem}
\label{thm:ut}
 The loop $L(G, T)$ terminates from every start state iff Definition~\ref{def:UT} is
 satisfiable.
\end{theorem}

\begin{proof}
 Then the codomain of $R$ is a well founded set and $R$ is an order homomorphism.

 (Do this proof properly).
\end{proof}

The existence of a ranking function is equivalent to the satisfiability
of the formula from Definition~\ref{def:UT}, so a witness to the satisfiability
of formula {\bf [UT]} is also a ranking function and thus a proof of $L$'s unconditional termination.

@@

\begin{theorem}
\label{thm:ont}
 The loop $L(G, T)$ has an infinite execution iff Defintion~\ref{def:ont} is satisfiable.
\end{theorem}

\begin{proof}
 Then $x_0$ is an underapproximation to the reachable states on entry to $L$, and $N$ is an open recurrence set.

 (Do this proof properly).
\end{proof}

@@

\begin{theorem}
\label{thm:cnt}
 If $T$ is deterministic, i.e. each state $x$ has exactly one successor $x'$ such that $T(x, x')$, then
 Definition~\ref{def:ont} is equivalent to Definition~\ref{def:cnt}.
\end{theorem}

\begin{proof}
 If $T$ is deterministic, then each $x$ has a unique successor $S(x)$, so $\forall x, x' . T(x, x') \leftrightarrow x' = S(x)$.

 (Finish this proof).
\end{proof}

@@

Conditional termination is then equivalent to the
formula {\bf [CT]} of Defintion~\ref{def:ct}.

\begin{theorem}
\label{thm:ct}
 The loop $L(G, T)$ terminates when started in any state satisfying $I(x)$ iff
 Defintion~\ref{def:ct} is satisfiable.
\end{theorem}

\begin{proof}
 Do this proof.
\end{proof}

If this formula is satisfiable, two witnesses are returned:
\begin{itemize}
\item $W$ is an inductive invariant of $L$ that is established by the initial states $I$ if the loop
guard $G$ is met.
\item $R$ is a ranking function for $L$ as restricted by $W$ -- that is to say, $R$ need only
be well founded on those states satisfying $W \wedge G$.  Since $W$ is an inductive invariant of $L$,
$R$ is strong enough to show that $L$ terminates from any of its initial states.
\end{itemize}

@@

\begin{proof}
 This is a direct consequence of Theorems~\ref{thm:ut}, \ref{thm:ct}, \ref{thm:ont}, \ref{thm:cnt} and \ref{thm:synth-sound}.
\end{proof}

@@

 \section{Conclusions and Related Work}
%\subsection{Termination analysis}
Automated program termination is a research topic that has received a fair amount of attention from the software verification community.
In order to compare our technique to the rest of the area, 
Figure~\ref{fig:handletable} summarises the related works with respect to the restrictions they impose on the transition relations as well as the form of the computed ranking functions.
